---
title: 'BUDA 525: Team 4 Final Project'
author: "Ryan Antonini, Danny Germain, Joshua Meadows, Josh Nelson, Bill Robertson"
date: "9/27/2019"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Problem 1



# Problem 2



# Problem 3

```{r, echo=FALSE}
library(ISLR)
library(effects)
library(car)
head(Credit)
summary(Credit)
```

```{r,echo=FALSE}
Credit$cardsF <-as.factor(Credit$Cards)
mod3_1 <- lm(Balance~ID+Income+Limit+Rating+cardsF+Education+Gender+Student+Married+Ethnicity,data=Credit)
#We start by fitting a model with all predictors to find out how to procede
summary(mod3_1)
```

```{r,echo=FALSE}
mod3_2 <- lm(Balance~ID+Income+Limit+Rating+cardsF+Education+Student+Married,data=Credit)
#Next we fit a model without Gender and Ethnicity since we don't want to use them, and compare them to the model with all predictors
summary(mod3_2)
anova(mod3_1,mod3_2)
```

```{r,echo=FALSE}
mod3_3 <- lm(Balance~Income+Limit+Rating+cardsF+Student,data=Credit)
summary(mod3_3)
anova(mod3_1,mod3_3)
#testing to see if we need Education and Married
```

```{r}
plot(mod3_3)
avPlots(mod3_3)
plot(allEffects(mod3_3))
#Running Diagnostics
```

There is some serious NCV that needs delt with, coming from entries where Balance=0

```{r}
Credit2<- Credit[Credit$Balance!=0,]
mod3_4 <- lm(Balance~Income+Limit+cardsF+Student,data=Credit2)
summary(mod3_4)
```

```{r}
plot(mod3_4)
avPlots(mod3_4)
plot(allEffects(mod3_4))
ncvTest(mod3_4)
#Running Diagnostics
```

By removing the 90 cases where balance is zero, we can do an extremely good job at prediciting balance with only 4 predictors. Next, we create a new variable that reports whether or not the person has a balance, and use this variable as a response in a Logistic Regression model.

```{r}
Credit3<-Credit
Credit3$BalanceF<- as.factor(Credit3$Balance>0)
mod3_5<- glm(BalanceF~Limit+Student+Rating+cardsF+Age+Education+Gender+Married+Ethnicity+Income+ID,family = "binomial",data=Credit3)
summary(mod3_5)
mod3_6<-stepAIC(mod3_5)
summary(mod3_6)
mod3_7<- glm(BalanceF~Limit+Student,family = "binomial",data=Credit3)
summary(mod3_7)
```

```{r}
Credit3$Predict <- ifelse(mod3_6$fitted.values >0.5,"TRUE","FALSE")
mytable <- table(Credit3$BalanceF,Credit3$Predict)
print(mytable)
efficiency <- sum(diag(mytable))/sum(mytable)
efficiency

Credit3$Predict <- ifelse(mod3_7$fitted.values >0.5,"TRUE","FALSE")
mytable <- table(Credit3$BalanceF,Credit3$Predict)
print(mytable)
efficiency <- sum(diag(mytable))/sum(mytable)
efficiency
```

While Model 3_6 uses the fewest predictors to still have 100% effeciency (meaning it predicts correctly in all cases), none of the individual coeffeicients is significant. For this reason, I prefer Model 3_7 which has 96% efficiency using only two predictors, each of which is significant. We can now use Model 3_7 to predict whether or not.

# Problem 4

```{r}
library(carData)
head(Salaries)
summary(Salaries)
```

```{r}
mod4_1<-lm(salary~rank+discipline+yrs.since.phd+yrs.service+sex, data=Salaries)
summary(mod4_1)
```

```{r}
plot(mod4_1)
avPlots(mod4_1)
plot(allEffects(mod4_1))
```
